---
title: "Factors Affecting Voting Behavoir"
subtitle: "Data Mining in R | Class Exercise 1"
author: "Ursulina Kölbener, Marco Graf und Sascha Eng"
date: "2023-03-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

[tbd: what this work is about]

```{r cars}
summary(cars)
```

## Methods

Y-variable: 
* `vote` - Voted last national election
  + 1	Yes
  + 2	No
  + 3	Not eligible to vote
  + 7	Refusal (missing value)
  + 8	Don't know (missing value)
  + 9	No answer (missing value)

X-Variable:
* `polintr` - How interested in politics
  + 1	Very interested
  + 2	Quite interested
  + 3	Hardly interested
  + 4	Not at all interested
  + 7	Refusal (missing value)
  + 8	Don't know (missing value)
  + 9	No answer (missing value)
* `gndr` - Gender
  + 1	Male
  + 2	Female
  + 9	No answer (missing value)
* `agea` - Age of respondent, calculated
  + age as numer
  + 999	Not available (missing value)
* `domicil` - Domicile, respondent's description
  + 1	A big city
  + 2	Suburbs or outskirts of big city
  + 3	Town or small city
  + 4	Country village
  + 5	Farm or home in countryside
  + 7	Refusal (missing value)
  + 8	Don't know (missing value)
  + 9	No answer (missing value)

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Results

## Discussion
