---
title: "Factors Affecting Voting Behavoir"
subtitle: "Data Mining in R | Class Exercise 1"
author: "Ursulina Kölbener, Marco Graf und Sascha Eng"
date: "2023-03-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

[tbd: what this work is about]

```{r}
library(tidyverse) # Load the tidyverse package

# import dataset
ess <- rio::import(here::here("data", "ESS10.csv"))

ess_1 <- ess %>% 
  select(idno, vote, polintr, gndr, agea, domicil, region, eisced, eduyrs, cntry, ctzcntr) %>%  # select variables
  filter(cntry == "CH") %>%  # restrict sample to Switzerland
  filter(ctzcntr == 1) %>% # restrict sample to Swiss citizens 
  filter(vote %in% c(1,2)) %>%  # restrict to "yes" and "no" votes
  filter(!polintr %in% c(7,8,9)) %>%  # exclude values 7, 8, and 9
  filter(!agea %in% c(16,17,999)) %>% # exclude ages 16 and 17 and value 999
  filter(!domicil %in% c(7,8,9)) %>%   # exclude values 7, 8, and 9
  filter(eduyrs %in% c(5:21)) %>% # restrict sample to 5 to 21 years of education
  # Create factor variables 
  mutate(vote = factor(vote, levels = c(1,2), labels = c("Yes","No"))) %>% 
  mutate(polintr = factor(polintr, levels = c(1,2,3,4), 
                          labels = c("Very interested",
                                     "Quite interested",
                                     "Hardly interested",
                                     "Not at all interested"))) %>% 
  mutate(gndr = factor(gndr, levels = c(1,2), labels = c("Male", "Female"))) %>% # create binary factor 'gndr'
  mutate(domicil = factor(domicil, levels = c(1:5), labels = c("1","2","3","4","5"))) %>%  # create binary factor 'domicil'
  mutate(domicil = fct_collapse(domicil, Urban = c(1,2,3), Rural = c(4,5))) %>%  # collapse categories
  mutate(region = recode_factor(as.factor(region),  # create unordered factor 'region'
                                CH01 = "Lake Geneva",
                                CH02 = "Middleland",
                                CH03 = "North-West",
                                CH04 = "Zurich",
                                CH05 = "Eastern Part",
                                CH06 = "Central Part",
                                CH07 = "Ticino")) %>% 
  mutate(eisced = factor(eisced, ordered = TRUE, labels = c("ES-ISCED I",  # create ordered factor and name it 'educ'
                                                            "ES-ISCED II",
                                                            "ES-ISCED IIIb",
                                                            "ES-ISCED IIIa",
                                                            "ES-ISCED IV",
                                                            "ES-ISCED V1",
                                                            "ES-ISCED V2"))) %>% 
  rename(educ = eisced) %>% 
  mutate(age_group = factor(agea, ordered = TRUE, levels = c(18:90))) %>%  # create ordered factor 'age_group'
  mutate(age_group = fct_collapse(age_group, young = c(18:35), middle = c(36:55), older = c(56:90))) # collapse categories
 
 
```

## Methods

Y-variable: 
* `vote` - Voted last national election
  + 1	Yes
  + 2	No
  + 3	Not eligible to vote
  + 7	Refusal (missing value)
  + 8	Don't know (missing value)
  + 9	No answer (missing value)

X-Variable:
* `polintr` - How interested in politics
  + 1	Very interested
  + 2	Quite interested
  + 3	Hardly interested
  + 4	Not at all interested
  + 7	Refusal (missing value)
  + 8	Don't know (missing value)
  + 9	No answer (missing value)
* `gndr` - Gender
  + 1	Male
  + 2	Female
  + 9	No answer (missing value)
* `agea` - Age of respondent, calculated
  + age as numer
  + 999	Not available (missing value)
* `domicil` - Domicile, respondent's description
  + 1	A big city
  + 2	Suburbs or outskirts of big city
  + 3	Town or small city
  + 4	Country village
  + 5	Farm or home in countryside
  + 7	Refusal (missing value)
  + 8	Don't know (missing value)
  + 9	No answer (missing value)

```{r}
# some summary statistics
summary(ess_1)
```


```{r pressure, echo=FALSE}
ggplot(data = ess_1) +
  geom_jitter(mapping = aes(vote, polintr))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Results

## Discussion
